Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-19 doc/GMP2015/s03_drp.tex

The following subsections provide the supporting proofs.

@@

\subsubsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
\label{sec:union_intersection}

@@

\subsubsection{Inferring Parent Structure from Union and Intersection of Children}
\label{sec:infer_parent}

@@

\subsection{Example}

@@

% \begin{example}
%     In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan. These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$. This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 1. Also shown is the decomposition of the individual doublets. These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them. As explained in Theorem \ref{theorem:main}, point 2, all of them must be included in the optimal DR-plan. Were any left out, the union of the children would not be the entire parent.
% \end{example}

@@

This section details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}. The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$. These intersect on $G$ without triangle $1$ and $N$ which is clearly well-constrained. As explained in the proof of Theorem \ref{theorem:main}, part 2, since there are only 2 possible children, their intersection must occur 2 levels below the parent. Just as expected, it is on the third level, as it is a child of both of $G$'s children. Furthermore, it intersects trivially with the edges that connect triangle $1$ and $N$ to $2$ and $N-1$, respectively; these necessarily intersect with triangle $1$ and $N$ themselves trivially (by Theorem \ref{theorem:main}, part 1) making all of these children. The triangles then decompose into their constituent $K_{3,3}$ graphs, which then decompose into 9 edges.

>>>>

% The following subsections provide the supporting proofs.

@@

% \subsubsection{Relationships for the Union and Intersection of Well-Constrained Subgraphs}
% \label{sec:union_intersection}

@@

% \subsubsection{Inferring Parent Structure from Union and Intersection of Children}
% \label{sec:infer_parent}

@@

@@

% \myexample
% In figure \ref{fig:doublets}, the two well-constrained vertex-maximal proper subgraphs of the graph are the two doublets, shown in the second level of the DR-plan. These intersect on the well-constrained graph $\{b_1,b_2,b_3\}$. This will necessarily be a child $N=2$ levels below the parent of the two doublets, as explained in Theorem \ref{theorem:main}, point 1. Also shown is the decomposition of the individual doublets. These have five well-constrained vertex-maximal proper subgraphs, the two triangles and the three edges connecting them. As explained in Theorem \ref{theorem:main}, point 2, all of them must be included in the optimal DR-plan. Were any left out, the union of the children would not be the entire parent.

@@

\myexample
\textsl{[DR-Plan for self-similar structure]}
This section details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}. The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$. These intersect on $G$ without triangle $1$ and $N$ which is clearly well-constrained. As explained in the proof of Theorem \ref{theorem:main}, part 2, since there are only 2 possible children, their intersection must be a node 2 levels below the parent. Just as expected, it is on the third level, as it is a child of both of $G$'s children. Furthermore, it intersects trivially with the edges that connect triangle $1$ and $N$ to $2$ and $N-1$, respectively; these necessarily intersect with triangle $1$ and $N$ themselves trivially (by Theorem \ref{theorem:main}, part 1) making all of these children. The triangles then decompose into their constituent $K_{3,3}$ graphs, which then decompose into 9 edges.

### 2014-12-19 doc/GMP2015/s04_recombination.tex

\medskip\noindent
\textbf{Example}. \todo{This uses different example format, not their environment. Should this, or section 3 change?}
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving
isometries), with distinct orientation types $\sigma_p$ (two
orientation choices for each of the 4 triangles) that can be obtained
by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
size 2). By subdivided binary search in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
$(G,\delta)$ is found when the length of the nonedge $e$ in  the
realization $q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

>>>>

\myexample
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving
isometries), with distinct orientation types $\sigma_p$ (two
orientation choices for each of the 4 triangles) that can be obtained
by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
size 2). By subdivided binary search in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
$(G,\delta)$ is found when the length of the nonedge $e$ in  the
realization $q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\cite{sitharam2011cayleyI,sitharam2011cayleyII} include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

### 2014-12-19 doc/GMP2015/s05_problems.tex

The \dfn{restricted optimal DR-planning problem} requires DR-plans of
one of these types, which reduces to recognizing if the input graph is
a 2-tree or a tree-decomposable graph for which simple near-linear
time algorithms are available \uncited and the DR-plan is a
by-product output of the recognition algorithm.

@@

\subsection{Optimal Modification, Completion and Recombination: formal
connections}
%
The OMD problem is closely related to a well-studied problem of
completion of an under-constrained system to an isostatic one with a
small DR-plan.
\begin{observation}
\label{obs:OC_to_OMB}
    The (decision version of) the {\bf optimal completion problem
    (OC)} from \cite{bib:survey, j-a, fang} is OMD$_0$.
\end{observation}
In fact, a \dfn{restricted OC} problem was studied by \cite{j-a}
requiring the completion to be tree-decomposable. 
%It is not clear if
%the reverse reduction holds, i.e, is OMD reducible to OC?

@@

%\note This section deals only with bar-joint systems, however,
%extensions of several of these problems, approaches and solutions to
%other qusecs beyond  bar-joint - specifically pinned line incidence
%systems and multi-body-pin systems will be expanded in Sections
% \ref{sec:pinnedline} and \ref{sec:bodypin}.

>>>>

The \dfn{restricted optimal DR-planning problem} requires DR-plans of
one of these types, which reduces to recognizing if the input graph is
a 2-tree or a tree-decomposable graph for which simple near-linear
time algorithms are available \cite{valdes1979recognition,fudos1997graph} and the DR-plan is a
by-product output of the recognition algorithm.

@@

%In the recombination setting, the corresponding \dfn{restricted
%OMD$_k$  problem} requires the reduced graph $G'$ and its isostatic
%completion $H$ to be 2-trees as in Section \ref{sec:2-tree-reduction}
%or to be a low Cayley complexity tree-decomposable graph as in Section
%\ref{sec:tdecomp}. Clearly these problems have deterministic
%polynomial time algorithms, the algorithms run in time exponential in
%$k$. As a result, when $k$ is part of the input, the \dfn{restricted
%OMD problem} has the potential to be difficult. For example, when the
%isostatic completion $H$ is required to be a 2-tree the restricted OMD
%problem is reducible to the maximum spanning series-parallel subgraph
%problem shown by \cite{cai1993spanning} to be NP-complete even if the input
%graph is planar of maximum degree at most 6. However, since the OMD
%problem has other input restrictions such as not having any proper
%isostatic subgraphs, it is not clear if the reverse reduction exists
%and hence it is unclear whether the OMD problem is NP-complete. The
%same holds for the restricted OMD problem where the isostatic
%completion $H$ is required to be a tree-decomposable graph of low
%Cayley complexity (i.e, have special, small DR-plans). One potential
%obstacle to an indecomposable graph $G$'s membership in the restricted
%OMD$_k$ for small $k$ is if $G$ is tri-connected and has large girth.
%In fact, 6-connected (hence rigid) graphs with arbitrarily large girth
%have been constructed in \cite{servatius2000rigidity}.
%
\subsection{Optimal Modification, Completion and Recombination: formal
connections}
%
The OMD problem is closely related to a well-studied problem of
completion of an under-constrained system to an isostatic one with a
small DR-plan.
\begin{observation}
    The (decision version of) the {\bf optimal completion problem
    (OC)} from \cite{sitharam2005combinatorial,joan-arinyo2003transforming,zhang-gao2006well} is OMD$_0$.
\end{observation}
In fact, a \dfn{restricted OC} problem was studied by \cite{joan-arinyo2003transforming}
requiring the completion to be tree-decomposable. 
%It is not clear if
%the reverse reduction holds, i.e, is OMD reducible to OC?

@@

%\note This section deals only with bar-joint systems, however,
%extensions of several of these problems, approaches and solutions to
%other qusecs beyond  bar-joint - specifically pinned line incidence
%systems and multi-body-pin systems will be expanded in Sections
%\ref{sec:pinnedline} and \ref{sec:bodypin}.

### 2014-12-19 doc/GMP2015/s08_pinnedline.tex

The optimal qusecs DR-plan can also be applied to analyze and solve
the structure of  cross-linking microfibrils in collagen and cellulose
in biology, by modeling these structures as \dfn{pinned line-incidence
systems}.

\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical property \cite{buehler2008nanomechanics}.
\dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}) \cite{fall2013physical,smith1971plant}.
Both of these substances consist of a large number \dfn{microfibrils},
%fibril itself straight?
%Each fibril is attached to some fixed larger organelle/membrane at one spot (typically on the boundary).
%It is additionally
each of which is cross-linked at 2 places with some two other fibrils,
where the \dfn{cross-linking} is like an incidence constraint that the
two crosslinked fibrils can slide with each other (see
Figure~\ref{fig:crosslink}).

@@

A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$
together with parameters $\delta$ specifying $|E|$ \textbf{pins} with
fixed positions in $\mathbb{R}^2$, such that each edge is constrained
to lie on a line passing through the corresponding pin, i.e.\ $\delta:
E \rightarrow \mathbb{R}^2$.
%
% $G=(V,E, w_V, w_E)$
%%is a \textbf{2-dimensional pinned line-incidence graph},
% has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
%Additionally, there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
%such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where
%for each edge $e \in E$, there is a uniquely corresponding pin $x \in X$,
%such that $e$ is constrained to lie on a line passing through $x$.
%%There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
%
A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$ \cite{sitharam2014incidence}. Note that no trivial motion exists since the pins have fixed positions on the plane.
In particular, both a single vertex and a single edge are under-constrained graphs.

@@

Figure~\ref{fig:pinned_line} shows an example of a pinned line-incidence graph,
where the grey ovals stand for pins representing attachments of fibrils,
and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
% and the black dots stand for vertices/cross-linkings.
%There are  and 10 vertices/cross-linkings in the graph.
The graph is well-constrained, with 12 vertices and 24 edges/pins.

@@

\begin{remark}\label{lem:pinned_union_intersection}
Let $F_i$ and $F_j$ to be subgraphs of the same well-constrained graph $F$,
where each of them can be either a single edge or a connected well-constrained subgraph.
There are only two possible cases:
(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are well-constrained, if and only if $F_i \cup F_j$ is well-constrained, if and only if $F_i \cap F_j$ is well-constrained.
%The following table summarizes all bijections:
%\setlength{\tabcolsep}{3pt}
%{\small
%\begin{center}
%\begin{tabular}{|p{2.8cm}|c|p{2.3cm}|}
%\hline
%\textbf{Property of $F_i,F_j$} & \textbf{$F_i\cup F_j$} & \centering \textbf{$F_i\cap F_j$}  \tabularnewline \hline
% \centering--- &  \sout{over-constrained}          &  \centering --- \tabularnewline \hline
%at least one of $F_i,F_j$ is an edge & under-constrained       &  \centering a vertex (under-constrained)  \tabularnewline \hline
%both $F_i$ and $F_j$ are well-constrained & well-constrained        &  \centering well-constrained  \tabularnewline \hline
% \centering --- & ---                     &  \centering \sout{under-constrained} \tabularnewline \hline
%\end{tabular}
%\end{center}
%}
%\setlength{\tabcolsep}{6pt}
\end{remark}

Given Remark~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned
line-incidence graphs is given in Appendix~\ref{sec:appendix_pinned}.
%The following lemma serves as Lemma~ for pinned line-incidence and can be proved similarly.
%\begin{lemma}
%\label{lem:vertex_intersection}
%If $C_i\cap C_j$ is a single vertex, then $\forall k: C_i\cap C_k$ is a single vertex.
%\end{lemma}

>>>>

The optimal DR-plan of Section \ref{sec:DRP} can also be applied to analyze and solve
the structure of  cross-linking collagen and cellulose microfibrils 
in animal and plant cell walls, and actin filaments in the cytoskeleton
by modeling these structures as \dfn{pinned line-incidence
systems}.

\dfn{Collagen} is an important protein material in biological tissues with highly elastic mechanical property \cite{buehler2008nanomechanics}.
\dfn{Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}) \cite{fall2013physical,smith1971plant}.
Both of these substances consist of a large number \dfn{microfibrils},
%fibril itself straight?
%Each fibril is attached to some fixed larger organelle/membrane at one spot (typically on the boundary).
%It is additionally
each of which is cross-linked at 2 places with usually 3 other fibrils,
where the \dfn{cross-linking} is like an incidence constraint that the
crosslinked fibrils can slide against each other while remaining incident (see
Figure~\ref{fig:crosslink}).

@@

A \dfn{pinned line-incidence system} $(G,\delta)$ is a graph $G=(V,E)$
together with parameters $\delta$ specifying $|E|$ \textbf{pins} with
fixed positions in $\mathbb{R}^2$, such that each edge is constrained
to lie on a line passing through the corresponding pin, i.e.\ $\delta:
E \rightarrow \mathbb{R}^2$.
%
% $G=(V,E, w_V, w_E)$
%%is a \textbf{2-dimensional pinned line-incidence graph},
% has weight functions $w_V(v) = 2, w_E(e) = 1$ for all $v \in V, e \in E$.
%Additionally, there is  a  set $X$ of $|E|$ \textbf{pins} with fixed positions in $\mathbb{R}^2$,
%such that  %every edge $e \in E$ corresponds to a pin $x \in X$ where
%for each edge $e \in E$, there is a uniquely corresponding pin $x \in X$,
%such that $e$ is constrained to lie on a line passing through $x$.
%%There exists a set $X$ of $|E|$ pins fixed in $\mathbb{R}^2$, such that each edge $e_i$
%
A pinned line-incidence graph $G$ is rigid if $|E| = 2|V|$ and $|E'| \le 2|V'|$ for every induced subgraph $(V',E')$ \cite{sitharam2014incidence}. Note that no trivial motion exists since the pins have fixed positions on the plane.
Euclidean transformations are not factored out.
In particular, both a single vertex and a single edge are under-constrained graphs.

@@

Figure~\ref{fig:pinned_line} shows an example of a pinned line-incidence graph,
where the grey ovals denote pins representing attachments of fibrils,
and the vertices $a_1,a_2,\ldots, c_3$ represent cross-linkings.
% and the black dots stand for vertices/cross-linkings.
%There are  and 10 vertices/cross-linkings in the graph.
The graph is well-constrained, with 12 vertices and 24 edges/pins.

@@

\begin{lemma}\label{lem:pinned_union_intersection}
Let $F_i$ and $F_j$ to be subgraphs of the same well-constrained graph $F$,
where each of them can be either a single edge or a connected well-constrained subgraph.
There are only two possible cases:
(1) At least one of $F_i$, $F_j$ is an edge, if and only if $F_i \cup F_j$ is under-constrained, if and only if $F_i \cap F_j$ is trivial; and (2) both $F_i$ and $F_j$ are well-constrained, if and only if $F_i \cup F_j$ is well-constrained, if and only if $F_i \cap F_j$ is well-constrained.
%The following table summarizes all bijections:
%\setlength{\tabcolsep}{3pt}
%{\small
%\begin{center}
%\begin{tabular}{|p{2.8cm}|c|p{2.3cm}|}
%\hline
%\textbf{Property of $F_i,F_j$} & \textbf{$F_i\cup F_j$} & \centering \textbf{$F_i\cap F_j$}  \tabularnewline \hline
% \centering--- &  \sout{over-constrained}          &  \centering --- \tabularnewline \hline
%at least one of $F_i,F_j$ is an edge & under-constrained       &  \centering a vertex (under-constrained)  \tabularnewline \hline
%both $F_i$ and $F_j$ are well-constrained & well-constrained        &  \centering well-constrained  \tabularnewline \hline
% \centering --- & ---                     &  \centering \sout{under-constrained} \tabularnewline \hline
%\end{tabular}
%\end{center}
%}
%\setlength{\tabcolsep}{6pt}
\end{lemma}

Given Lemma~\ref{lem:pinned_union_intersection}, % and the modified definition of DR-plan,
proofs of Lemma~\ref{lemma:wc_intersection_is_C} and \ref{lemma:uc_intersection_makes_all_uc} straightforwardly extend to  pinned line-incidence graphs.
The proof of Lemma~\ref{lemma:wc_intersection_makes_all_wc} for pinned
line-incidence graphs is given in Appendix~\ref{sec:appendix_pinned}.
%The following lemma serves as Lemma~ for pinned line-incidence and can be proved similarly.
%\begin{lemma}
%\label{lem:vertex_intersection}
%If $C_i\cap C_j$ is a single vertex, then $\forall k: C_i\cap C_k$ is a single vertex.
%\end{lemma}
